Last glacial cycle glacier erosion potential in the Alps

### 2021-01-25 articles/erosion/alps-erosion.tex
Rename uncertainty figure as powerlaws figure.
Reshuffle erosion discussion part. Seeing clearer.
Draft choice of erosion law. A mess. Call it 0.4.

    Modelled erosion rates, $\dot{e}$, are calculated from the modelled basal
    velocities, $u_\mathrm{b}$, using a empirical erosion power-law,
    %
    \begin{equation}
        \dot{e} = K_\mathrm{g} u_\mathrm{b}^l ,
    \end{equation}
    %
    where $K_g = 5.2\times 10^{-11}\,m^{1-l}\,a^{l-1}$ and $l = 2.34$ are
    empirical constants calibrated on quantified sediment yields and velocities
    from 15~Patagonian and Antarctic Peninsula tidewater glaciers
    \citep{Koppes.etal.2015}. It is important to note that there are both large
    uncertainties surrounding these numbers on the aforementioned glaciers
    \todo{quantify}, and little insurance that this erosion law applies
    elsewhere. Alternative erosion laws are analyzed in the discussion part
    (Sect.~\ref{sec:uncertainty}).

@@

% -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -
\subsection{Results interpretation}
% -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -

@@

@@

% -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -
\subsection{Sensitivity to climate}
\label{sec:sensitivity}
% -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -

    The modelled erosion potential depends significantly on the choice of
    pleoclimate forcing applied, but the general patterns remain similar.
    Higher precipitation yields higher erosion. The results also appear to be
    sensitive to model resolution, primarily in the mountains interior, where
    discretized slopes depend on the grid size.

    Fig.~\ref{fig:sensitivity} -- Sensitivity to climate.

% -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -
\subsection{Erosion uncertainties}
\label{sec:uncertainty}
% -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -

    The cumulative erosion potential modelled by the Franz-Joseph Glacier
    erosion law \citep{Herman.etal.2015} appear overestimated. However, the
    error bars are large.

    The erosion law by \citet{Cook.etal.2020} results in similarly high values
    but a flatter (less localized) pattern of cumulative erosion potential,
    with lower values in the troughs and more erosion in the mountains. The two
    erosion laws may appear incompatible, and there would be a lot of reasons
    for them to be incompatible. But as pointed by \citet{Cook.etal.2020}, they
    are not necessarily incompatible. The erosion's feedback onto ice dynamics
    may yield to a delocalization of high-velocity flow in the long run.

    The erosion law by \citet{Koppes.etal.2015} yields a similar pattern but
    lower values. This law is for tidewater glaciers, which flow much faster
    than mountain glaciers. However, we can imagine that Alpine paleoglaciers
    flowing on a thick sediment bed (and sometimes into proglacial lakes, not
    included), behave closer to tidewater glaciers than the Franz-Josef steep
    mountain glacier.

    Fig.~\ref{fig:uncertainty} -- Erosion uncertainties

@@

    \begin{figure*}
      \centerline{\includegraphics{alpero_uncertainty}}
      \caption{%
        Modelled cumulative (time-integrated) glacial erosion potential for
        three different erosion laws published by
        \textbf{(a)} \citet{Cook.etal.2020},
          $1.665 \times 10^{-1} \textbf{u}_\mathrm{b} ^{0.6459}$,
        \textbf{(b)} \citet{Herman.etal.2015},
          $2.7 \times 10^{-7} \textbf{u}_\mathrm{b} ^{2.02}$
          (same as Fig.~\ref{fig:cumulative}a), amd
        \textbf{(c)} \citet{Koppes.etal.2015},
          $5.2 \times 10^{-8} \textbf{u}_\mathrm{b} ^{2.34}$.
        \textbf{(d)} Corresponding erosion power laws, and
        \textbf{(e)} modelled cumulative (time-integrated) glacial erosion
          potential along a Rhine Glacier transect (upper panels dashed line).}
      \label{fig:uncertainty}
    \end{figure*}

>>>>

    Modelled erosion rates, $\dot{e}$, are calculated from the modelled basal
    velocities, $u_\mathrm{b}$, using a empirical erosion power-law,
    %
    \begin{equation}
        \dot{e} = K_\mathrm{g} u_\mathrm{b}^l ,
    \end{equation}
    %
    where $K_g = 5.2\times 10^{-11}\,m^{1-l}\,a^{l-1}$ and $l = 2.34$ are
    empirical constants calibrated on quantified sediment yields and velocities
    from 15~Patagonian and Antarctic Peninsula tidewater glaciers
    \citep{Koppes.etal.2015}. It is important to note that there are both large
    uncertainties surrounding these numbers on the aforementioned glaciers
    \todo{quantify}, and little insurance that this erosion law applies
    elsewhere. Alternative erosion laws are analyzed in the discussion part
    (Sect.~\ref{sec:powerlaws}).

@@

% -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -
\subsection{Climate sensitivity}
\label{sec:sensitivity}
% -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -

    \note{I feel there is not much to learn from this section. I am thinking
          about just leaving it out.}

    The modelled erosion potential depends significantly on the choice of
    paleoclimate forcing applied, but the general patterns remain similar.
    Higher precipitation yields higher erosion. The results also appear to be
    sensitive to model resolution, primarily in the mountains interior, where
    bedrock slopes and sling speed depend more strongly on the grid size
    (Fig.~\ref{fig:sensitivity}).

% -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -
\subsection{Choice of erosion law}
\label{sec:powerlaws}
% -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -

    The choice of erosion law significantly impacts results presented in this
    study (Fig.~\ref{fig:powerlaws}). The erosion law based on quantified
    sediment yields from Patagonian and Antarctic Peninsula tidewater glaciers
    \citet[$\dot{e} = 5.2 \times 10^{-8} u_\mathrm{b}^{2.34}$]{Koppes.etal.2015},
    our default choice, yields moderate yet strongly localized cumulative erosion
    potential (Figs~\ref{fig:cululative} and Fig.~\ref{fig:powerlaws}a).

    The erosion law based on measured suspended sediment load and composition
    in the outlet creek of Franz-Joseph Glacier
    \citep[$\dot{e} = 2.7 \times 10^{-7} u_\mathrm{b}^{2.02}$]{Herman.etal.2015},
    also yields strongly localized erosion potential but much higher values
    (Fig.~\ref{fig:powerlaws}b).

    The erosion law based on a global compilation of glacier velocity and erosion
    rates
    \citet[$\dot{e} = 1.665 \times 10^{-1} u_\mathrm{b}^{0.6459}$]{Cook.etal.2020},
    results in similarly high values, but a flatter (less localized) pattern of
    cumulative erosion potential (Fig.~\ref{fig:powerlaws}ac), with lower
    values in the troughs and more erosion in the mountains.

    These three erosion laws appear incompatible. Given the scarcity of measured
    glacier erosion rates, and the variety in glacier systems, climate and
    geologic setting, this is not a surprise.  But as more extensively
    discussed by \citet{Cook.etal.2020}, they do not have to be entirely
    incompatible. The erosion's feedback onto ice dynamics
    may yield to a delocalization of high-velocity flow in the long run.

    Alternative empirical erosion laws by
    \citet{Herman.etal.2015, Cook.etal.2020} yield glacial cycle cumulative
    erosion potential several orders of magnitude higher than modelled with the
    law by \citet{Koppes.etal.2015}, with values in the kilometres or tens of
    kilometres in places, which is of course well above the existing glacial
    relief from the Alps resulting from multiple glacial cycles.

    It is surprising that the erosion law by \citet{Koppes.etal.2015},
    calibrated on tidewater glaciers, yields more realistic erosion rates for
    the glacial Alps. However, we can imagine that Alpine paleoglaciers
    flowing on a thick sediment bed (and sometimes into proglacial lakes, not
    included), behave closer to tidewater glaciers than the Franz-Josef steep
    mountain glacier. Moreover, the cold climate of the LGM Alps resembled more
    that of modern Patagonia and Antarctica, than that of Franz-Joseph Glacier
    or most glaciers in the global compilation by \citet{Cook.etal.2020}.

% -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -
\subsection{Deglaciation rapid erosion}
% -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -

@@

% -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -
\subsection{Age of the glacial landscape}
% -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -

@@

@@

    \begin{figure*}
      \centerline{\includegraphics{alpero_powerlaws}}
      \caption{%
        Modelled cumulative (time-integrated) glacial erosion potential for
        three different erosion laws published by
        \textbf{(a)} \citet{Koppes.etal.2015},
          $5.2 \times 10^{-8} u_\mathrm{b} ^{2.34}$ (same as
          Fig.~\ref{fig:cumulative}a but with a different colour scale),
        \textbf{(b)} \citet{Herman.etal.2015},
          $2.7 \times 10^{-7} u_\mathrm{b} ^{2.02}$, and
        \textbf{(c)} \citet{Cook.etal.2020},
          $1.665 \times 10^{-1} u_\mathrm{b} ^{0.6459}$.
        \textbf{(d)} Corresponding erosion power laws, and
        \textbf{(e)} modelled cumulative (time-integrated) glacial erosion
          potential along a Rhine Glacier transect (upper panels dashed line).}
      \label{fig:powerlaws}
    \end{figure*}
